\documentclass[11pt,numbers,nocopyrightspace,acmlarge,anonymous]{acmart}


\usepackage{amsmath,bm,amssymb,stmaryrd}
\usepackage{amsthm}
\usepackage{listings}

\lstset{ %
morekeywords={*,forall,Lam,lam,int}
}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

%\titlebanner{Under Review}        % These are ignored unless
%\preprintfooter{FunTAL: Reasonably mixing a functional language with assembly}   % 'preprint' option specified.

\title{Theorems for Free for Free: Parametricity, With and Without Types}
\subtitle{\vspace{-3cm}} % Yay double blindness, so much room!
          \onecolumn
          \begin{centering}
            \LARGE{Theorems for Free for Free: Parametricity, With and Without Types}
          \end{centering}
          \vspace{1cm}

\section{Abstract}
The polymorphic blame calculus integrates static typing, including
universal types, with dynamic typing. The primary challenge with this
integration is preserving parametricity: even dynamically-typed code
should satisfy it once it has been cast to a universal type.  Ahmed
et~al.~(2011) employ runtime type generation in the polymorphic blame
calculus to preserve parametricity, but a proof that it does so has
been elusive.  Matthews and Ahmed~(2008) gave a proof of parametricity
for a closely related system that combines ML and Scheme, but later
found a flaw in their proof.
%
In this paper we present an improved version of the polymorphic blame
calculus and we prove that it satisfies relational parametricity. The
proof relies on a step-indexed Kripke logical relation. The
step-indexing is required to make the logical relation well-defined in
the case for the dynamic type. The possible worlds include the mapping
of generated type names to their types and the mapping of type names
to relations. We prove the Fundamental Property of this logical
relation and that it is sound with respect to contextual equivalence.
%
To demonstrate the utility of parametricity in the polymorphic blame
calculus, we derive two free theorems.

\section{Artifact description}
We present a type checker and stepper for the Polymorphic Blame Calculus.
Programs written in the code box below may be typechecked and, if they pass,
loaded into the stepper. Programs in the stepper are split into redex and context
and may be evaluated either one step at a time or 100 steps at a time.
The global name store is also displayed.
Additionally, our artifact includes a step backwards button that reverses the last step taken.

\section{Getting Started}

The artifact is presented as an online editor / stepper at\\

\verb|http://www.ccs.neu.edu/home/dijamner/paramblame/artifact/| \\

\noindent The source code, which is linked from that page, is at\\

\verb|https://github.com/dijamner/paramblame| 

\section{Building artifact}

To build the artifact and run its test suite, the following options
are available:

\paragraph{OCaml} With \verb|OCaml 4.03| and \verb|opam| installed,
you can set the package up with \verb|make install-deps|, \verb|make|,
and then running \verb|./test.native|. Note that this has been tested
on Fedorah 22 and Mac OSX 10.11. To run the editor locally,
after building open the file \verb|artifact/index.html| in a browser.

\paragraph{Docker} With Docker (tested with 17.03.1-ce), you can build and
run the test suite by running 
\verb|docker build -t paramblame .| 
and then, once that finishes, \verb|docker run paramblame|, which will run the test
suite. If you'd like to run the editor locally, you must copy the
javascript out of the docker image with

\verb|docker run --rm paramblame cat /paramblame/artifact/web.js > artifact/web.js| 

\noindent and then open \verb|artifact/index.html| in a browser.

In both cases, you can then change the tests (in \verb|test.ml|) and
re-run the commands to run the modified tests (in the \verb|OCaml|
case, you need not run \verb|make install-deps| again, though it does
not hurt).

\section{Assessing artifact}

Our artifact should be evaluated on its usefulness
as an aid to help readers understand 
the operational semantics of the polymorphic blame calculus.
Thanks to the design of the artifact,
it allows readers to observe firsthand the
conversion generation procedures and 
runtime name generation that preserves parametricity in
this calculus.

This artifact will constitute a useful supplement
to our paper if it allows readers to follow along
with the examples in the paper and observe when
our language produces type errors,
when programs step to blame,
and when programs return results.
The best way to utilize this artifact is to
follow along with the examples in the paper,
as described below.

\section{Examples}

Our test suite is primarily composed of examples from
the paper. We have included a couple examples from the paper,
in the web page itself 
as well as a couple of additional examples,
to demonstrate how to write code for this artifact. 

\begin{itemize}
\item \textbf{Simple call} 
	A simple example from STLC with a function call
	to see how the artifact works.
\item \textbf{Omega} We use the dynamic type to 
	enable self-application and produce a diverging program.
\item \textbf{Factorial} We include factorial
	as an example of a program with meaningful computational
	content that uses the dynamic type.
\item \textbf{Invalid cast} This example demonstrates
	that our type checker does not permit casts
	that cannot work.
\item \textbf{Trying to use dynamic type without a cast}
	This example shows that our language uses casts
	as the only way to produce values of the dynamic type.
\end{itemize}

As a demonstration of the features of the artifact, enter the
following program into the code editor:

\begin{lstlisting}
((Lam X.lam (x : X).x) : forall X. X -> X => int -> int) 5
\end{lstlisting}

This program first casts the polymorphic identity function to a
function from \textbf{int} to \textbf{int} and then applies it to
the integer 5. Hit the "Type check \& load" button to run the type checker
and initialize the interpreter. The program should appear in the
interpreter below, broken into redex and context.

Since \textbf{int} to \textbf{int} is a specialization
of the type of the identity function, the cast should succeed and
this program should produce 5. Hit the "step $-\!\!>$" button to 
proceed to the next evaluation step. In this case, the cast breaks
down into an instantiation and a simpler cast.

Notice that the step counter to the left of the button has increased.
This counter serves as a convenience to help keep track of program execution.
Hit the "step $<\!\!-$" button to go back to the last step and see that 
the step counter decreases again.

Now step forward twice so that the counter reads "2". When the 
type abstraction is instantiated, three things happen: a new type name,
"a!0", is generated and bound to the dynamic type in the store 
at the bottom of the interpreter,
X is replaced with a!0 in the body of the abstraction, 
and the result is wrapped in a conversion from a function at type a!0 
to a function at dynamic type. Note that if this step is repeated,
it will generate a new name each time. While this is not strictly
necessary, it is sufficient to guarantee name freshness and simplifies
the implementation.

Hit the "step $-\!\!>$" button again. The value 5 is pushed under the
cast. This casts it to the dynamic type so the inner function may be applied
and then casts the result of that application back to \textbf{int}.

On the next step, the value moves under the conversion the same way
and is wrapped again. Notice that the conversion on the 5 has a negative
label. 

Step 5 applies the identity function, leaving the value 5 with
a series of casts and conversions. The next two steps cancel the 
conversions and the casts respectively and leave the 5 in the result box.

To see what happens when a cast fails, enter the following modified program:

\begin{lstlisting}
((Lam X.lam (x:X).(3 : int => * : * => X)) 
	: forall X. X -> X => int -> int) 5
\end{lstlisting}

This program attempts to violate parametricity by returning the constant 3
from a supposedly polymorphic function. As it should, this results in blame.

Step to step 4. The steps up to this point are identical to the previous code.
Take the next step. This calls the function and returns the cast 3.
Since X was replaced with a type name rather than the dynamic type,
the two casts are incompatible. Take another step; the term reduces to blame.
In one more step, the blame propagates up through the context.


\section{Limitations}
Users may only write programs under a closed store.
We feel that this represents the class of programs
that users are likely to write, since conversions
are best thought of as an internal implementation detail.
\end{document}
